Sound and Automated Synthesis of Digital Stabilizing Controllers for Continuous Plants \thanks{Supported by EPSRC grant EP/J012564/1 and ERC project 280053 (CPROVER).}

### 2017-02-01 hscc2017/paper.tex
work on text

\begin{mylemma}\cite{astrom1997computer}
%
Given a synchronized ZOH input and sample and hold output on the plant with
a sample time $T$ satisfying the Nyquist criterion, the discrete pulse
transfer function $G(z,T)$ is an exact z-domain representation of $G(s)$
that can be computed using the following formula:
%
\begin{equation}
\label{eq:pulsetf}
G(z,T) = %\mathcal{ZOH}\left\lbrace{G(s)}\right\rbrace =
(1-z^{-1})\mathcal{Z}\left\lbrace{\mathcal{L}^{-1}\left\lbrace{\frac{G(s)}{s}}\right\rbrace_{t=kT}}\right\rbrace.
\end{equation}
%
\end{mylemma}
%
For the sake of brevity, we will use the notation $G(z)$ to represent the
pulse transfer function.

Lemma~\ref{eq:pulsetf} ensures that the poles and zeros match under the
$\mathcal{Z}\left\lbrace{\mathcal{L}^{-1}\left\lbrace{\cdot}\right\rbrace_{t=kT}}\right\rbrace$
operations, and it includes the ZOH dynamics in the $(1-z^{-1})$ term, which
is sufficient for stability studies over $G(s)$.

@@

CEGIS-based control synthesis requires a formal verifier to check whether a
candidate controller meets the requirements when combined with the plant. 
We use the Digital-System Verifier (DSVerifier)~\cite{IsmailBCFF15} for this
purpose.  It checks the stability of closed-loop control systems and
considers finite-word length (FWL) effects in the digital controller and
uncertainty parameters in the plant model (plant intervals)~\cite{Bessa16}. 
Let $C(z)$ be a digital controller and $G(z)$ be a plant model given as
%
\begin{align}
\small
\label{controller_plant_tf}
C(z)&=\frac{C_n(z)}{C_d(z)}=\frac{\beta_{0}+\beta_{1}z^{-1}+...+\beta_{M_C}z^{-M_C}}{\alpha_{0}+\alpha_{1}z^{-1}+...+\alpha_{N_C}z^{-N_C}}, \\
%\label{plant_tf}
G(z)&=\frac{G_n(z)}{G_d(z)}=\frac{b_{0}+b_{1}z^{-1}+...+b_{M_G}z^{-M_G}}{a_{0}+a_{1}z^{-1}+...+a_{N_G}z^{-N_G}}.
\end{align}
%
\noindent where $\vec{\beta}$ and $\vec{\alpha}$ are vectors containing the
controller's coefficients; similarly, $\vec{b}$ and $\vec{a}$ denote the
plant's coefficients; and $N_{(\cdot)}$ and $M_{(\cdot)}$ indicate the order
of the polynomial of the corresponding block.

Uncertainties in $G(z)$ may appear due to uncertainties in $G(s)$ (i.e.,
$\Delta_p{G}_n(s),\Delta_p{G}_d(s)$, which arise from tolerances/imprecision
in the original model), errors in the numeric calculations, or roundoffs in
the quantization.  These uncertainties are parametrically expressed by
additive terms, resulting in an uncertain model $\hat{G}(z)$, such that:
%
\begin{equation}
\hat{G}(z)=\frac{G_n(z)+\Delta G_n(z)}{G_d(z)+\Delta G_d(z)},
\end{equation}
%
which will be represented by the following transfer function:
%
\begin{equation}
\hat{G}(z)=\frac{\hat{b}_{0}+\hat{b}_{1}z^{-1}+...+\hat{b}_{M_G}z^{-M_G}}{\hat{a}_{0}+\hat{a}_{1}z^{-1}+...+\hat{a}_{N_G}z^{-N_G}} : N_G \geq M_G .
\end{equation}
%
Due to the nature of the methods we use for the stability check, we require
that the parametric errors in the plant have the same polynomial order as
the plant itself (all other errors described in this paper are ensured to
fulfill this property in our implementation).

We employ a notation based on the polynomials coefficients to simplify 
the presentation.  Let $\mathcal{P}^{N}$ be a polynomial
domain of order $N$. Let $P \in \mathcal{P}^{M,N}$ be a rational polynomial
$\frac{P_n}{P_d}$, where $P_n \in \mathcal{P}^{M}$ and $P_d \in \mathcal{P}^{N}$.  
For a vector of coefficients
%
\begin{equation}
\vec{P} \in \mathbb{R}^{N+M+2}=[n_{0}\ n_{1}\ \hdots \ n_{M}\ d_{0}\ d_{1}\ \hdots\ d_{N}\ ]^T
\label{eq:coefficients}
\end{equation}
%
and an uncertainty vector 
%
\begin{equation}
\Delta{\vec{P}}=[\Delta{n}_{0}\ \Delta{n}_{1}\ \hdots \ \Delta{n}_{M}\ \Delta{d}_{0}\ \Delta{d}_{1}\ \hdots\ \Delta{d}_{N}\ ]^T \; 
\label{eq:delta_coefficients}
\end{equation}
%
we write $\vec{\hat{G}}=\vec{G}+\Delta{\vec{G}}$.  In the following we will
either use the functions $G(z)$, $C(z)$ directly, or work over their
respective coefficients $\vec{G}$, $\vec{C}$ in vector form.

@@

A typical digital control system with a continuous plant and a discrete
controller is illustrated in Figure~\ref{fig:sampledsystem}.  The DAC and
ADC converters introduce quantization errors which are modeled as noises
$\nu_{1}(z)$ and $\nu_{2}(z)$, $G(s)$ is the continuous-time plant model
with parametric additive uncertainty $\Delta_p{G}_n(s)$ and
$\Delta_p{G}_d(s)$, $R(z)$ is the reference signal, $U(z)$ is the control
signal, and $\hat{Y}(z)$ is the output signal affected by the noises and
uncertainties in the closed-loop system.
%
%
%
The ADC and DAC may be abstracted by transforming the closed-loop system in
Figure~\ref{fig:sampledsystem} into the digital system in 
Figure~\ref{fig:digsystem1}, where the parametric uncertainties are
represented by~$\Delta_p \vec{G}$
%
%=\mathcal{Z}\left\lbrace{G_{P}(s)}\right\rbrace$
%
and the entire effect of $\nu_{1}$ and $\nu_{2}$ in the output $Y(z)$ is
represented as an additive output noise~$\nu(z)$. 
%, thus
%$\hat{Y}(z)=Y(z)+\nu(z)$.  The measurement noise $\nu(k)$ is enough for
%represent both quantization noises ($\nu_{1}$ and $\nu_{2}$) what can be
%obtained reorganizing the equations~\ref{eq:outputfunctions}:
%%
%\begin{equation}
%\label{eq:outputfunctions2}
%\hat{Y}(z)=H_{1}R(z)+H_{1}\nu_{1}(z)+H_{2}(z)\nu_{2}(z),
%\end{equation}
%%
%where $H_{1}R(z)$ is exactly equal to $Y(z)$ without quantization noise effects, thus:
%%
%\begin{equation}
%\nu(z)=H_{1}\nu_{1}(z)+H_{2}(z)\nu_{2}(z).
%\end{equation}

@@

%  \path[->] (v1) edge (q1.north);
%  \path[->] (v2) edge (q2.north);
  \path[->] (r) edge (circle.west);
  \path[->] (nu) edge (circle2.north);
  \path[->] (aux1) edge (yz);
  \path  
   (circle.east) edge (cz)
   (cz.east) edge (gs.west)
   (gs.east) edge (circle2.west)
   (circle2.east) edge (aux1.west)
   (aux1.south) edge (aux2.north)
   (aux2.west) edge (aux3.east); 
  \path[->]  (aux3.north) edge (circle.south);
 \end{tikzpicture}
}
\caption{System equivalent to the system in Figure~\ref{fig:sampledsystem} 
\label{fig:digsystem1}}
\end{figure}

%
In Figure~\ref{fig:digsystem1}, two sources of uncertainty are illustrated:
parametric uncertainties related to modeling errors, and uncertainties
introduced by quantizations in sampling and holding process, i.e., ADC and
DAC conversions.  These sources are assumed to be non-deterministic, 
which poses no further restriction to our model, thus, the uncertain model
may be rewritten as a vector of coefficients in the z-domain using equations
(\ref{eq:coefficients}) and (\ref{eq:delta_coefficients})
%
\begin{equation}
\label{eq:complete_unc_model}
\vec{\hat{G}}=\vec{G}+\Delta_p \vec{G}+\Delta_q \vec{G},
\end{equation}
%
where $\Delta_p \vec{G}$ is the term that corresponds to the parametric
uncertainties and $\Delta_q \vec{G}$ is the term that reflects the
quantization effects on the plant model, {\it i.e.} $\Delta_q \vec{G}$ is 
a model that incorporates the effects of $\nu$ in $\vec{\hat{G}}$ to 
investigate the consequences in the stability of system, once our stability 
analysis is based on polynomial coefficients and not over signals.

>>>>

\begin{mylemma}\cite{astrom1997computer}
%
Given a synchronized ZOH input and sample-and-hold output on the plant with
a sample time $T$ satisfying the Nyquist criterion, the discrete pulse
transfer function $G(z,T)$ is an exact z-domain representation of $G(s)$
that can be computed using the following formula:
%
\begin{equation}
\label{eq:pulsetf}
G(z,T) = %\mathcal{ZOH}\left\lbrace{G(s)}\right\rbrace =
(1-z^{-1})\mathcal{Z}\left\lbrace{\mathcal{L}^{-1}\left\lbrace{\frac{G(s)}{s}}\right\rbrace_{t=kT}}\right\rbrace.
\end{equation}
%
\end{mylemma}
%
For the sake of brevity, we will use the notation $G(z)$ to represent the
pulse transfer function $G(z,T)$.

Lemma~\ref{eq:pulsetf} ensures that the poles and zeros match under the
$\mathcal{Z}\left\lbrace{\mathcal{L}^{-1}\left\lbrace{\cdot}\right\rbrace_{t=kT}}\right\rbrace$
operations, and it includes the ZOH dynamics in the $(1-z^{-1})$ term, which
is sufficient for stability studies over $G(s)$ \red{[we need to clarify the last sentence - why is it sufficient, where is this fact stated?]}.

@@

CEGIS-based control synthesis requires a formal verifier to check whether a
candidate controller meets the requirements when combined with the plant. 
We use the Digital-System Verifier (DSVerifier)~\cite{IsmailBCFF15} for this
purpose.  It checks the stability of closed-loop control systems and
considers finite-word length (FWL) effects in the digital controller, and
uncertainty parameters in the plant model (plant intervals)~\cite{Bessa16}. 
Let $C(z)$ be a digital controller and $G(z)$ be a plant model given as
%
\begin{align}
\small
\label{controller_plant_tf}
C(z)&=\frac{C_n(z)}{C_d(z)}=\frac{\beta_{0}+\beta_{1}z^{-1}+...+\beta_{M_C}z^{-M_C}}{\alpha_{0}+\alpha_{1}z^{-1}+...+\alpha_{N_C}z^{-N_C}}, \\
%\label{plant_tf}
G(z)&=\frac{G_n(z)}{G_d(z)}=\frac{b_{0}+b_{1}z^{-1}+...+b_{M_G}z^{-M_G}}{a_{0}+a_{1}z^{-1}+...+a_{N_G}z^{-N_G}}.
\end{align}
%
\noindent where $\vec{\beta}$ and $\vec{\alpha}$ are vectors containing the
controller's coefficients; similarly, $\vec{b}$ and $\vec{a}$ denote the
plant's coefficients; and $N_{(\cdot)}$ and $M_{(\cdot)}$ indicate the order
of the polynomials. 
% of the corresponding block.

Uncertainties in $G(z)$ may appear due to uncertainties in $G(s)$ (i.e.,
$\Delta_p{G}_n(s),\Delta_p{G}_d(s)$, which arise from tolerances/imprecision
in the original model), errors in the numeric calculations, or roundoffs in
the quantization.  These uncertainties are parametrically expressed by
additive terms, resulting in an uncertain model $\hat{G}(z)$, such that:
%
\begin{equation}
\hat{G}(z)=\frac{G_n(z)+\Delta G_n(z)}{G_d(z)+\Delta G_d(z)},
\end{equation}
%
which will be represented by the following transfer function:
%
\begin{equation}
\hat{G}(z)=\frac{\hat{b}_{0}+\hat{b}_{1}z^{-1}+...+\hat{b}_{M_G}z^{-M_G}}{\hat{a}_{0}+\hat{a}_{1}z^{-1}+...+\hat{a}_{N_G}z^{-N_G}} : N_G \geq M_G .
\end{equation}
%
\red{[why is this condition on the order of the polynomials raised only now? how is it used later?]}
Due to the nature of the methods we use for the stability check, we require
that the parametric errors in the plant have the same polynomial order as
the plant itself (\blue{all other errors described in this paper are ensured to
fulfil this property in our implementation}\red{[have we later argued that this is indeed the case?]}) 
\blue{and that enter the plant linearly}.

We employ a notation based on the polynomials coefficients to simplify 
the presentation.  Let $\mathcal{P}^{N}$ be the space of polynomials of order $N$. 
Let $P \in \mathcal{P}^{M,N}$ be a rational polynomial 
$\frac{P_n}{P_d}$, where $P_n \in \mathcal{P}^{M}$ and $P_d \in \mathcal{P}^{N}$.  
For a vector of coefficients
%
\begin{equation}
\vec{P} \in \mathbb{R}^{N+M+2}=[n_{0}\ n_{1}\ \hdots \ n_{M}\ d_{0}\ d_{1}\ \hdots\ d_{N}\ ]^T
\label{eq:coefficients}
\end{equation}
%
and an uncertainty vector 
%
\begin{equation}
\Delta{\vec{P}}=[\Delta{n}_{0}\ \Delta{n}_{1}\ \hdots \ \Delta{n}_{M}\ \Delta{d}_{0}\ \Delta{d}_{1}\ \hdots\ \Delta{d}_{N}\ ]^T \; 
\label{eq:delta_coefficients}
\end{equation}
%
we write $\vec{\hat{G}}=\vec{G}+\Delta{\vec{G}}$.  
\red{[confusing: we mix notations on abstract polynomials with those used for the plant above - the orders of the two are different ($N$ vs $N_G$) - please use the latter.]}
In the following we will
either use the functions $G(z)$, $C(z)$ directly, or work over their
respective coefficients $\vec{G}$, $\vec{C}$ in vector form.

@@

A typical digital control system with a continuous plant and a discrete
controller is illustrated in Figure~\ref{fig:sampledsystem}.  The DAC and
ADC converters introduce quantization errors which are modeled as noises
$\nu_{1}(z)$ and $\nu_{2}(z)$, $G(s)$ is the continuous-time plant model
with parametric additive uncertainty $\Delta_p{G}_n(s)$ and
$\Delta_p{G}_d(s)$, $R(z)$ is the reference signal, $U(z)$ is the control
signal, and $\hat{Y}(z)$ is the output signal affected by the noises and
uncertainties in the closed-loop system.
%
%
%
The ADC and DAC may be abstracted by transforming the closed-loop system in
Figure~\ref{fig:sampledsystem} into the digital system in 
Figure~\ref{fig:digsystem1}, where the parametric uncertainties are
represented by~$\Delta_p \vec{G}$
%
%=\mathcal{Z}\left\lbrace{G_{P}(s)}\right\rbrace$
%
and the entire effect of $\nu_{1}$ and $\nu_{2}$ in the output $Y(z)$ is
represented as an additive output noise~$\nu(z)$.  
\red{[We have recast the model as in Figure 2, but never defined $\nu$ and never use it later - retain the two separate noise contributions as in Figure 1.]}
%, thus
%$\hat{Y}(z)=Y(z)+\nu(z)$.  The measurement noise $\nu(k)$ is enough for
%represent both quantization noises ($\nu_{1}$ and $\nu_{2}$) what can be
%obtained reorganizing the equations~\ref{eq:outputfunctions}:
%%
%\begin{equation}
%\label{eq:outputfunctions2}
%\hat{Y}(z)=H_{1}R(z)+H_{1}\nu_{1}(z)+H_{2}(z)\nu_{2}(z),
%\end{equation}
%%
%where $H_{1}R(z)$ is exactly equal to $Y(z)$ without quantization noise effects, thus:
%%
%\begin{equation}
%\nu(z)=H_{1}\nu_{1}(z)+H_{2}(z)\nu_{2}(z).
%\end{equation}

@@

%  \path[->] (v1) edge (q1.north);
%  \path[->] (v2) edge (q2.north);
  \path[->] (r) edge (circle.west);
  \path[->] (nu) edge (circle2.north);
  \path[->] (aux1) edge (yz);
  \path  
   (circle.east) edge (cz)
   (cz.east) edge (gs.west)
   (gs.east) edge (circle2.west)
   (circle2.east) edge (aux1.west)
   (aux1.south) edge (aux2.north)
   (aux2.west) edge (aux3.east); 
  \path[->]  (aux3.north) edge (circle.south);
 \end{tikzpicture}
}
\caption{Digital equivalent to the system in Figure~\ref{fig:sampledsystem} 
\label{fig:digsystem1}}
\end{figure}

\red{[why is the output in Figure 2 tilde, whereas that in  figure 1 hat?]}
%
In Figure~\ref{fig:digsystem1}, two sources of uncertainty are illustrated:
parametric uncertainties related to modeling errors, and uncertainties
introduced by quantizations in sampling and holding process, i.e., ADC and
DAC conversions.  These sources are assumed to be non-deterministic, 
which poses no further restriction to our model, thus, the uncertain model
may be rewritten as a vector of coefficients in the z-domain using equations
(\ref{eq:coefficients}) and (\ref{eq:delta_coefficients})
%
\begin{equation}
\label{eq:complete_unc_model}
\vec{\hat{G}}=\vec{G}+\Delta_p \vec{G}+\Delta_q \vec{G},
\end{equation}
%
where $\Delta_p \vec{G}$ is the term that corresponds to the parametric
uncertainties and $\Delta_q \vec{G}$ is the term that reflects the
quantization effects on the plant model, {\it i.e.} $\Delta_q \vec{G}$ is 
a model that incorporates the effects of $\nu$ in $\vec{\hat{G}}$ to 
investigate the consequences in the stability of system, once our stability 
analysis is based on polynomial coefficients and not over signals.
